---
title: 'Project 1: Exploratory Data Analysis'
author: "SDS348"
date: ''
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, fig.align = "center", warning = F, message = F,
tidy=TRUE, tidy.opts=list(width.cutoff=60), R.options=list(max.print=100))
```

## Data Wrangling and Data Exploration

### Instructions
A knitted R Markdown document (ideally HTML) and the raw R Markdown file (as .Rmd) should both be submitted to Canvas by 11:59pm on 10/11/2020. These two documents will be graded jointly, so they must be consistent (i.e., don’t change the R Markdown file without also updating the knitted document).

The text of the document should provide a narrative structure around your code/output. All results presented must have corresponding code. Any answers/results/plots etc. given without the corresponding R code that generated the result will not be considered. Furthermore, all code contained in your final project document must work correctly (knit early, knit often)! Please do not include any extraneous code or code which produces error messages. (Code that produces warnings is acceptable, as long as you understand what the warnings mean!)

### Find data:

Find two (!) datasets with one variable in common (e.g., dates, times, states, counties, countries, sports players), both with at least 50 observations (i.e., rows) in each. Please think very carefully about whether it makes sense to combine your datasets! If you find one dataset with 50 patients and it has their age, and you find another dataset with 50 *different* patients that has their ages, it makes no sense to join them based on age (you would just be pairing up random people of the same age).

When combined, the resulting/final dataset must have **at least 4 different variables (at least 3 numeric) in addition to the common variable** (i.e., five variables total).

You can have as many variables as you would like! If you found two datasets that you like but they don't have enough variables, find a third dataset with the same common variable and join all three.



### Guidelines

1. If the datasets are not tidy, you will need to reshape them so that every observation has its own row and every variable its own column. If the datasets are both already tidy, you will make them untidy with `pivot_wider()/spread()` and then tidy them again with `pivot_longer/gather()` to demonstrate your use of the functions. It's fine to wait until you have your descriptives to use these functions (e.g., you might want to pivot_wider() to rearrange the data to make your descriptive statistics easier to look at); it's fine long as you use them at least once!

    - Depending on your datasets, it might be a good idea to do this before joining. For example, if you have a dataset you like with multiple measurements per year, but you want to join by year, you could average over your numeric variables to get means/year, do counts for your categoricals to get a counts/year, etc.
    
    - If your data sets are already tidy, demonstrate the use of `pivot_longer()/gather()` and `pivot_wider()/spread()` on all or part of your data at some point in this document (e.g., after you have generated summary statistics in part 3, make a table of them wide instead of long).
    

2. Join your 2+ separate data sources into a single dataset based on a common ID variable! If you can't find a good pair datasets to join, you may split one main dataset into two different datasets with a common ID variable in each, and then join them back together based on that common ID, but this is obviously less than ideal.

    - You will document the type of join that you do (left/right/inner/full), including a discussion of how many observations were in each dataset, which observations in each dataset were dropped (if any) and why you chose this particular join. 


3. Create summary statistics

    - Use *all six* core `dplyr` functions (`filter, select, arrange, group_by, mutate, summarize`) to manipulate and explore your dataset. For mutate, create a  new variable that is a function of at least one other variable, preferably using a dplyr vector function (see dplyr cheatsheet). It's totally fine to use the `_if`, `_at`, `_all` versions of mutate/summarize instead (indeed, it is encouraged if you have lots of variables)
    
    - Create summary statistics (`mean, sd, var, n, quantile, min, max, n_distinct, cor`, etc) for each of your numeric variables both overall and after grouping by one of your categorical variables (either together or one-at-a-time; if you have two categorical variables, try to include at least one statistic based on a grouping of two categorical variables simultaneously). If you do not have any categorical variables, create one using mutate (e.g., with `case_when` or `ifelse`) to satisfy the `group_by` requirements above. Ideally, you will find a way to show these summary statistics in an easy-to-read table (e.g., by reshaping). (You might explore the kable package for making pretty tables!) If you have lots of numeric variables (e.g., 10+), or your categorical variables have too many categories, just pick a few (either numeric variables or categories of a categorical variable) and summarize based on those. It would be a good idea to show a correlation matrix for your numeric variables (you will need it to make one of your plots).
 
4. Make visualizations (three plots)

    -  Make a correlation heatmap of your numeric variables
    -  Create at least two additional plots of your choice with ggplot that highlight some of the more interesting features of your data.
    - Each plot (besides the heatmap) should have at least three variables mapped to separate aesthetics
    - Each should use different geoms (don't do two geom_bars)
    - At least one plot should include `stat="summary"`
    - Each plot should include a supporting paragraph describing the relationships that are being visualized and any trends that are apparent
        - It is fine to include more, but limit yourself to 4. Plots should avoid being redundant! Four bad plots will get a lower grade than two good plots, all else being equal.
    - Make them pretty! Use correct labels, etc.
    
    
5. Perform k-means/PAM clustering or PCA on (at least) your numeric variables.

    - Include all steps as we discuss in class, including a visualization.

    - If you don't have at least 3 numeric variables, or you want to cluster based on categorical variables too, convert them to factors in R, generate Gower's dissimilarity matrix on the data, and do PAM clustering on the dissimilarities.
    
    - Show how you chose the final number of clusters/principal components 
    
    - Interpret the final clusters/principal components 

    - For every step, document what your code does (in words) and what you see in the data!     

<P style="page-break-before: always">
\newpage
    
### Rubric

Prerequisite: Finding appropriate data from at least two sources per the instructions above: Failure to do this will result in a 0! You will submit a .Rmd file and a knitted document (pdf).

#### 0. Introduction (4  pts)

- Write a narrative introductory paragraph or two describing the datasets you have chosen, the variables they contain, how they were acquired, and why they are interesting to you. Expand on potential associations you may expect, if any.

<br/>
I chose two datasets (acquired as .csv files), both from a website called Kaggle relating to music artists. One dataset contains information about grammy award winners since 1960 as well as the categories they were nominated for, the year they won, as well as the project they were nominated for. The other dataset I chose details the various numerical details relating to that artist in terms of plays in a particular music data website (lastfm). Both of these datasets paint an illustration of award winning artists and their aggregated number of plays and listeners, and I was interested in finding the differences between Grammy winners in terms of the year they won as well as the category they won for, and the impact of this on their overall success (assessed through the amount of interaction their music received in this particular website). I estimate that newer artists and artists that have won more prestigious Grammy awards such as Best Album or Best Song will do much better numbers-wise, overall. Also, I estimate that the majority of high yielding numbers will originate from artists that are based in the United States, as the Grammys are usually awarded to artists within the country.

#### 1. Tidying: Rearranging Wide/Long (8 pts)

- Tidy the datasets (using the `tidyr` functions `pivot_longer`/`gather` and/or `pivot_wider`/`spread`) 
- If you data sets are already tidy, be sure to use those functions somewhere else in your project
- Document the process (describe in words what was done per the instructions)

```{r}
#obtain libraries
library(readxl)
library(writexl)
library(dplyr)
library(gdata)
library(tidyverse)

#obtain process
grammys = read_csv("the_grammy_awards.csv")
grammys = grammys %>% select(-c(title,published_at,workers,updated_at,img,winner))
artists = read_csv("artists.csv")
#remove blank nominee cells
grammys <- grammys[!(is.na(grammys$artist) | grammys$artist==""), ]
```

<br/>
I read in both datasets which are .csv files, and removed the columns that I thought were unnecessary. Note that I removed the unnecessary columns from the artists dataset a little later in the process, so you'll see that in the next step after they're joined. These datasets were already clean, as I already had a plan in mind when I chose this project to combine by artist, which was already set up as a column in both files. The only cleaning I really had to do was remove cells with missing information, as well as remove some unnecessary columns, but that was done after the merge had already occurred.
    
#### 2. Joining/Merging (8 pts)

- Join your datasets into one using a `dplyr` join function
- If you have multiple observations on the joining variable in either dataset, fix this by collapsing via summarize
- Discuss the process in words, including why you chose the join you did
- Discuss which cases were dropped, if any, and potential problems with this

```{r}
#set up joining
grammys <- grammys %>% mutate(join_title=str_replace_all(str_trim(str_to_lower(artist)), fixed(" "), ""))
head(grammys)
artists <- artists %>% mutate(join_title=str_replace_all(str_trim(str_to_lower(artist_lastfm)), fixed(" "), ""))
head(artists)
merged <- grammys %>% inner_join(artists)
head(merged)
merged <- merged %>% select(-c(mbid, join_title, artist_mb, artist_lastfm, country_lastfm))
merged <- merged %>% mutate(country_mb = ifelse(str_length(country_mb)==0, 'Unknown', as.character(country_mb)))
filtered_countries <- merged %>% filter(country_mb!='Unknown')
unique_artists <- merged %>% select(-c(category, nominee, year)) %>% distinct()

# pivoting
mergedtag <- unique_artists %>% filter(tags_lastfm!='') %>% mutate(tag = strsplit(as.character(tags_lastfm), ';'))
mergedtag <- mergedtag %>% mutate(tag = as.character(lapply(tag, `[[`, 1)))
pivot_merged <- mergedtag %>% pivot_wider(names_from = tag, values_from = scrobbles_lastfm)
head(pivot_merged)
```

<br/>
To prepare for joining, I created a separate column that took the artist names column in that dataset, lowercased it, and then removed all whitespaces. I did this primarily so that it would be more inclusive for potential joins that wouldn't occur due to some formatting difference between the two datasets. I opted to use an inner-join to combine the two datasets as not having one or the other left out important data that would be pointless since I couldn't gain any insight from columns that didn't provide any relationships between the two dataset. Also, the artists dataset contains an incredible amount of artists, and since I was only focused on Grammy-winning artists, the majority of artists in the dataset  weren't going to be extremely relevant to my analysis. Therefore, all artists that didn't appear in both datasets were dropped (~2000 in grammys, and ~1.2 mil in artists). There were probably some artists that were dropped falsely as there were some weird formatting issues in some artist names that created special characters, but other than that, there weren't really any problems created from this join step other than a relatively small loss of potential useful information.

#### 3. Wrangling (40 pts)

- Use all six core `dplyr` functions in the service of generating summary statistics (18 pts)
    - Use mutate at least once to generate a variable that is a function of at least one other variable
    
```{r}

#six dplyr
#added new numeric column averaging scrobble per listener
merged <- merged %>% mutate(scrobble_per_listener = scrobbles_lastfm/listeners_lastfm)
#alphabetic numeric values of artists
merged %>% group_by(artist) %>% select_if(is.numeric) %>% arrange(artist)
#list of all artists in alphabetical order
merged %>% group_by(artist) %>% select(artist) %>% arrange(artist)
#mean scrobbles of artists that have won a Grammy for Album of the Year by country of origin
merged %>% filter(category == "Album Of The Year") %>% group_by(country_mb) %>% summarize(mean_scrobbles = mean(scrobbles_lastfm))
#mean listeners of artists by genre of music
mergedtag %>% filter(!is.na(tag)) %>% group_by(tag) %>% summarize(mean = mean(listeners_lastfm)) %>% as.data.frame

```
    

- Compute at least 10 summary statistics for using summarize and summarize with group_by (18 pts)
    - Use at least 5 unique functions inside of summarize (e.g., mean, sd)
    - At least 2 of these should group by a categorical variable. Create one by dichotomizing a numeric if necessary
    - If applicable, at least 1 of these should group by two categorical variables
    - Strongly encouraged to create a correlation matrix with `cor()` on your numeric variables
    
```{r}
# summary statistics
stat1 <- merged %>% select(listeners_lastfm) %>% 
  summarize(mean = mean(listeners_lastfm), sd = sd(listeners_lastfm), 
            var = var(listeners_lastfm), n = n(), min = min(listeners_lastfm),
            max = max(listeners_lastfm), n_dist = n_distinct(listeners_lastfm))

stat2 <- merged %>% select(scrobbles_lastfm) %>% 
  summarize(mean = mean(scrobbles_lastfm), sd = sd(scrobbles_lastfm), 
            var = var(scrobbles_lastfm), n = n(), min = min(scrobbles_lastfm),
            max = max(scrobbles_lastfm), n_dist = n_distinct(scrobbles_lastfm))

stat3 <- merged %>% select(scrobble_per_listener) %>% filter(!is.na(.)) %>%
  summarize(mean = mean(scrobble_per_listener), sd = sd(scrobble_per_listener), 
            var = var(scrobble_per_listener), n = n(), min = min(scrobble_per_listener),
            max = max(scrobble_per_listener), n_dist = n_distinct(scrobble_per_listener))

stat4 <- merged %>% select(year) %>% 
  summarize(mean = mean(year), sd = sd(year), 
            var = var(year), n = n(), min = min(year),
            max = max(year), n_dist = n_distinct(year))

#grouped summary statistics
stat5 <- merged %>% group_by(country_mb, artist) %>% select(listeners_lastfm) %>% 
  summarize(mean_lstnrs = mean(listeners_lastfm), sd_lstnrs = sd(listeners_lastfm), 
            var_lstnrs = var(listeners_lastfm), n_lstnrs = n(), min_lstnrs = min(listeners_lastfm),
            max_lstnrs = max(listeners_lastfm), n_dist_lstnrs = n_distinct(listeners_lastfm))

stat6 <- merged %>% group_by(country_mb) %>% select(scrobbles_lastfm) %>% 
  summarize(mean_scrobbles = mean(scrobbles_lastfm), sd_scrobbles = sd(scrobbles_lastfm), 
            var_scrobbles = var(scrobbles_lastfm), n_scrobbles = n(), min_scrobbles = min(scrobbles_lastfm),
            max_scrobbles = max(scrobbles_lastfm), n_dist_scrobbles = n_distinct(scrobbles_lastfm))

stat7 <- merged %>% group_by(country_mb) %>% select(scrobble_per_listener) %>%
  summarize(mean_scrob_per_listen = mean(scrobble_per_listener), sd_scrob_per_listen = sd(scrobble_per_listener), 
            var_scrob_per_listen = var(scrobble_per_listener), n_scrob_per_listen = n(), min_scrob_per_listen = min(scrobble_per_listener),
            max_scrob_per_listen = max(scrobble_per_listener), n_dist_scrob_per_listen = n_distinct(scrobble_per_listener))

stat8 <- merged %>% group_by(country_mb) %>% select(year) %>% 
  summarize(mean_scrob_per_listen = mean(year), sd_scrob_per_listen = sd(year), 
            var_scrob_per_listen = var(year), n_scrob_per_listen = n(), min_scrob_per_listen = min(year),
            max_scrob_per_listen = max(year), n_dist_scrob_per_listen = n_distinct(year))

stat9 <- merged %>% group_by(country_mb, artist) %>% select(scrobbles_lastfm) %>% 
  summarize(mean_scrob = mean(scrobbles_lastfm), sd_scrob = sd(scrobbles_lastfm), 
            var_scrob = var(scrobbles_lastfm), n = n(), min_scrob = min(scrobbles_lastfm),
            max_scrob = max(scrobbles_lastfm), n_dist_scrob = n_distinct(scrobbles_lastfm))

stat10 <- merged %>% group_by(country_mb, artist) %>% select(year) %>% 
  summarize(mean_yr = mean(year), sd_yr = sd(year), 
            var_yr = var(year), n_yr = n(), min_yr = min(year),
            max_yr = max(year), n_dist_yr = n_distinct(year))

#quantiles
quantile(merged$listeners_lastfm)
quantile(merged$scrobbles_lastfm)
quantile(merged$scrobble_per_listener, na.rm=T)
quantile(merged$year)

#create table
combined_stats <- rbind(stat1, stat2, stat3, stat4)
combined_country_stats <- cbind(stat6, stat7, stat8)
combined_grouped_stats <- cbind(stat5, stat9, stat10)
#truncated columns to 5
knitr::kable(head(combined_stats[, 1:5]), "simple", row_names=c("stat1", "stat2", "stat3", "stat4"))
#truncated columns to 5
knitr::kable(head(combined_grouped_stats[, 1:5]), "pipe")
#truncated columns to 5
knitr::kable(head(combined_country_stats[, 1:5]), "pipe")
```


- Summarize/discuss all results in no more than two paragraphs (4 pts)

The first four stats relate to the simple summary statistics (mean, sd, var, n, min, max, n_distinct) of all four of the numeric variables in my data, these being scrobbles, listeners, scrobbles per listeners, and year. For year, we see what we expect to see, the numerical summaries of all possible years with some slight skew due to an overabundance of information on some years (mostly later ones) than others. For the other three numeric measurement statistics, there was mostly expected data, although through combing through these statistics, I realized that there were some missing collections for some of the artists, as min for scrobbles and listeners were 0, indicating that for some artists, this data was not collected at all. I also noticed that the scrobbles per artist was the higest variation category exhibiting standard deviation in the 10s of millions. Also, there were some NAs obtained through the process of computing them with my scrobble and listeners data so the count of these columns were not as high. One last thing I noticed was that the mean of scrobbles was higher than the mean amount of listeners by two magnitudes which indicates that there were listeners that listened to these artists more than once.

For the more complicated summary statistics that involved grouping, I noticed that there was some pretty high variation of datas between countries, and there were some pretty surprising countries that were top of particular columns (like UK for highest mean scrobbles). I noticed that my data in the U.K. was particularly highly varied as it contained 0 as the minimum amount of scrobbles for an artist as well as the maximum amount of scrobbles too (coming from The Beatles). Unsurprisingly, I noticed that the majority of artists that there was data for were from the United States, since most Grammy winners are from the U.S. I also noticed that in my data, the higest amount of Grammy rows I had for a grouping of artist with a country was U2 who are from Ireland, although after eliminating some NAs from my scrobble_per_listener column, it became Arethra Franklin. Also, note that the correlation matrix for all numeric variables is calculated in the next step, not in this step.


#### 4. Visualizing (30 pts)

- Create a correlation heatmap of your numeric variables

```{r}
#remove NA from numeric data processed in plotting
merged_no_na <- merged %>% na.omit
#display correlation matrix for numeric variables in dataset
merged_no_na %>% select_if(is.numeric) %>% cor()
#obtain color pallette
my_colors <- RColorBrewer::brewer.pal(6, "YlOrRd")[1:6]
merged_no_na%>%select_if(is.numeric)%>%cor%>%as.data.frame%>%
  rownames_to_column%>%pivot_longer(-1)%>%  ggplot(aes(rowname,name,fill=value))+
  geom_tile()+  geom_text(aes(label=round(value,2)))+  xlab("")+ylab("")+coord_fixed()+
  scale_fill_gradient2(low="red",mid=my_colors[1],high=my_colors[4]) + 
  theme(axis.text.x = element_text(hjust=0.5, angle = 9))
```

The correlation matrix details the relationship between all numeric variables in the dataset. By looking at the heatmap above,
we can see that the columns, scrobbles and listeners seem to be heavily correlated (value of 0.85), thus being shaded the darkest, and
the columns, year and listeners, are the least correlated with a value of simply 0.1. There are no negative correlation values in the
dataset.

- Create two effective, polished plots with ggplot

    - Each plot should map 3+ variables to aesthetics 
    - Each plot should have a title and clean labeling for all mappings
    - Change at least one default theme element and color for at least one mapping per plot
    - For at least one plot, add more tick marks (x, y, or both) than are given by default
    - For at least one plot, use the stat="summary" function
    - Supporting paragraph or two (for each plot) describing the relationships/trends that are apparent
    
```{r}
merged <- merged %>% mutate(older = (year < 2000))
#rename country names to isocode for cleaner graph
library(countrycode)
filtered_iso = filtered_countries %>% mutate(country_mb = countrycode(country_mb, origin='country.name', destination='iso2c'))

category_numbers <- merged %>% group_by(category) %>% mutate(total_listeners = sum(listeners_lastfm), total_scrobbles = sum(scrobbles_lastfm),
count = n(), year = as.factor(year)) %>% distinct(category, .keep_all=T)

#plot
library(ggplot2)
library(RColorBrewer)
library(scales)
ggplot(merged, aes(x= artist, y = listeners_lastfm, color = older)) + geom_point(size=2) +  ylab("listeners") + 
  theme(axis.text.x = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(color = "Artist Won Before 2000") + scale_color_brewer(palette = "Dark2") +
  scale_y_discrete(label=comma, limits = c(600000, 1200000, 1800000, 2400000, 3000000, 3600000, 4200000, 4800000)) + labs(title="Artists plotted to # of listeners")

```
    
When I first thought up of combining this data, I was expecting this graph to heavily favor
the newer winners of the Grammys in terms of the interaction they obtain from a website that aggregates listening information like last.fm. What I hypothesized is that there would not be as many people listening to older music, and therefore creating a noticeable difference between the listeners that an older Grammy-winning artist had compared to a newer Grammy-winning artist. However, after looking at the scatterplot, this does not seem to be the case, as the graph does not seem to show any correlation between these two categories of how long ago the artist won the Grammys compared to how many listeners they are recorded to have in last.fm.

```{r}
ggplot(filtered_iso, aes(country_mb, fill = country_mb)) +
  stat_summary(aes(y=scrobbles_lastfm), fun = "mean", geom = "bar", color="black") + ylab("average # of plays per listener") + 
  xlab("countries of origin for grammy winners") + labs(fill = "country") +
  theme(legend.position="none", panel.background = element_rect(fill = "grey"), plot.title = element_text(hjust = 0.5)) + scale_y_continuous(labels = function(x) format(x, scientific = F)) + labs(title = "Average # of plays of artists per originating country")

```

By looking at the plot for this data, I noticed some things that I expected to see, and some things that I didn't. The United States, as I expected, rated extremely well for this category as most Grammy winners from the U.S. tend to have high commercial success. However by looking at the data, we can see that the United States is not the country with the highest number of plays, but rather the United Kingdom. After analyzing the data, it appears that this is due to large variation in the data, and outliers pushing the numbers of the United Kingdom up (particularly the Beatles). Also, some surprising countries that I didn't expect to rate so highly in this category but did, include Ireland and Romania.
    
#### 5. Dimensionality Reduction (20 pts) 

- Either k-means/PAM clustering or PCA (inclusive "or") should be performed on at least three numeric variables in your dataset

    - All relevant steps discussed in class 
    - A visualization of the clusters or the first few principal components (using ggplot2)
    - Supporting paragraph or two describing results found 

```{r}
#clustering
library(cluster)
clust_dat<-merged_no_na%>%select(scrobbles_lastfm, scrobble_per_listener, listeners_lastfm)

#determine number of clusters
sil_width <- vector()
for(i in 2:10){ 
  kms <- kmeans(clust_dat, centers = i)
  sil <- silhouette(kms$cluster, dist(clust_dat))
  sil_width[i] <- mean(sil[, 3])
}
ggplot() + geom_line(aes(x = 1:10, y = sil_width)) + scale_x_continuous(name = "k", breaks = 1:10)

#determine that 2 is best cluster number
kmeans <- clust_dat %>% scale %>% kmeans(2)
kmeansclust<-clust_dat%>%mutate(cluster=as.factor(kmeans$cluster))
kmeansclust%>%ggplot(aes(scrobbles_lastfm,listeners_lastfm,color=cluster))+geom_point()
library("plot3D")
col_list <- c("red", "blue")
scrobbles <- kmeansclust$scrobbles_lastfm
scrobble_per_listener <- kmeansclust$scrobble_per_listener
listeners <- kmeansclust$listeners_lastfm
colors <- kmeansclust$cluster
scatter3D(scrobbles, scrobble_per_listener, listeners, colvar = NULL, col = col_list[colors],
          pch = 19, cex = 0.5)
```

I used k-means clustering on three of the numeric variables (scrobbles, listeners, scrobbles per listeners) in my dataset. By using the silhouette method, I determined that there should be 2 clusters for this particular data, as this is where the silhouette width is the higest. This indicated that there were two groups that each datapoint in my dataset could identify as. By using the native R function kmeans, I clustered my three numeric variables into these two clusters, and displayed it in a 2d plot representation mapping scrobbles to listeners. However, the 2d graph may cause confusion in why these data points are clustered this way as it does not display all three dimensions that were assessed to determine the clusters, therefore, I included a 3d plot to better display that insight as well.

#### 6. Neatness!

- Your project should not knit to more than 30 or so pages (probably closer to 10)! You will lose points if you print out your entire dataset(s), have terrible formatting, etc. If you start your project in a fresh .Rmd file, you are advised to paste the set-up code from this document (lines 14-17) at the top of it: this will automatically truncate if you accidentally print out a huge dataset, etc. Imagine this is a polished report you are giving to your PI or boss to summarize your work researching a topic.


### Where do I find data?

OK, brace yourself!

You can choose ANY datasets you want that meet the above criteria for variables and observations. I'm just sitting here but off the top of my head, if you are into amusement parks, you could look at amusement-park variables, including ticket sales per day etc.; then you could join this by date in weather data. If you are interested in Game of Thrones, you could look at how the frequency of mentions of character names (plus other character variables) and the frequency of baby names in the USA...You could even take your old Biostats data and merge in new data (e.g., based on a Google forms timestamp).

You could engage in some "me-search": You can [request your Spotify data](https://support.spotify.com/ca-en/article/data-rights-and-privacy-settings/) or [download Netflix viewing activity](https://help.netflix.com/en/node/101917), Amazon purchase history, etc. You can use your Google Fit/Fitbit/Apple watch data, etc. These can be combined (e.g., with each other, with other data sources).

You can make it as serious as you want, or not, but keep in mind that you will be incorporating this project into a portfolio webpage for your final in this course, so choose something that really reflects who you are, or something that you feel will advance you in the direction you hope to move career-wise, or something that you think is really neat. On the flip side, regardless of what you pick, you will be performing all the same tasks, so it doesn't end up being that big of a deal.

If you are totally clueless and have no direction at all, log into the server and type 

```{R eval=F}
data(package = .packages(all.available = TRUE))
```

This will print out a list of **ALL datasets in ALL packages** installed on the server (a ton)! Scroll until your eyes bleed! Actually, do not scroll that much... To start with something more manageable, just run the command on your own computer, or just run `data()` to bring up the datasets in your current environment. To read more about a dataset, do `?packagename::datasetname`. 

If it is easier for you, and in case you don't have many packages installed, a list of R datasets from a few common packages (also downloadable in CSV format) is given at the following website: https://vincentarelbundock.github.io/Rdatasets/datasets.html (including types/numbers of variables in each)

- A good package to download for fun/relevant data is `fivethiryeight`. Just run `install.packages("fivethirtyeight"), load the packages with `library(fivethirtyeight)`, run `data()`, and then scroll down to view the datasets. Here is an online list of all 127 datasets (with links to the 538 articles). Lots of sports, politics, current events, etc: https://cran.r-project.org/web/packages/fivethirtyeight/vignettes/fivethirtyeight.html

- If you have already started to specialize (e.g., ecology, epidemiology) you might look at discipline-specific R packages (vegan, epi, respectively). We will be using some tools from these packages later in the course, but they come with lots of data too, which you can explore according to the directions above

- However, you *emphatically DO NOT* have to use datasets available via R packages! In fact, I would much prefer it if you found the data from completely separate sources and brought them together (a much more realistic experience in the real world)! You can even reuse data from your SDS328M project, provided it shares a variable in common with other data which allows you to  merge the two together (e.g., if you still had the timestamp, you could look up the weather that day: https://www.wunderground.com/history/). If you work in a research lab or have access to old data, you could potentially merge it with new data from your lab!

- Here is a curated list of interesting datasets (read-only spreadsheet format): https://docs.google.com/spreadsheets/d/1wZhPLMCHKJvwOkP4juclhjFgqIY8fQFMemwKL2c64vk/edit

- Here is another great compilation of datasets: https://github.com/rfordatascience/tidytuesday

- Here is the UCI Machine Learning Repository: https://archive.ics.uci.edu/ml/index.php

    - See also https://en.wikipedia.org/wiki/List_of_datasets_for_machine-learning_research#Biological_data

- Here is another good general place to look: https://www.kaggle.com/datasets

- To help narrow your search down or to see interesting variable ideas, check out https://www.tylervigen.com/spurious-correlations. This is the spurious correlations website, and it is fun, but if you look at the bottom of each plot you will see sources for the data. This is a good place to find very general data (or at least get a sense of where you can scrape data together from)!

- If you are interested in medical data, check out www.countyhealthrankings.org

- If you are interested in scraping UT data, the university makes *loads* of data public (e.g., beyond just professor CVs and syllabi). Check out all the data that is available in the statistical handbooks: https://reports.utexas.edu/statistical-handbook

##### Broader data sources:

[Data.gov](www.data.gov) 186,000+ datasets!

[Social Explorer](Social Explorer) is a nice interface to Census and American Community Survey data (more user-friendly than the government sites). May need to sign up for a free trial.

[U.S. Bureau of Labor Statistics](www.bls.gov)

[U.S. Census Bureau](www.census.gov)

[Gapminder](www.gapminder.org/data), data about the world.

...





